---
title: "Nicfall_originalhwcode_02"
author: "Nicola Kriefall"
date: "9/26/2019"
output: html_document
---

#Notes on the hw subject
##Poisson Distribution
The Poisson Distribution is often used to model open ended counts of independently occuring events, for example the number of cars that pass a traffic intersection over a given interval of time or the number of times a monkey scratches itself during a given observation interval. The probability mass function for the Poisson distribution is described by a single parameter, λ, where λ can be interpreted as the mean number of occurrences of the event in the given interval.

Let’s use R to look at the probability mass functions for different values of λ:
```{r}
x <- 0:10
l = 3.5
probset<-dpois(x=x,lambda=l)
barplot(probset,names.arg=x,space=0,xlab="x",ylab="Pr(X = x)",main="Probability Mass Function")
```

Cumulative function distribution:
```{r}
x <- 0:10
l <- 3.5
barplot(ppois(q=x,lambda=l),ylim=0:1,space=0, names.arg=x,xlab="x",ylab="Pr(X ≤ x)",main="Cumulative Probability")
```

#Homework Problem
Every Saturday, at the same time, a primatologist goes and sits in the forest in the morning and listens for titi monkey calls, counting the number of calls they hear in a 2 hour window from 5am to 7am. Based on previous knowledge, she believes that the mean number calls she will hear in that time is exactly 15. Let X represent the appropriate Poisson random variable of the number of calls heard in each monitoring session.

What is the probability that she will hear more than 8 calls during any given session?
```{r}
p <- ppois(q=8,lambda=15) #probability that she will hear up to & including 8 calls/session
q <- 1-p #probability that she will hear more than 8 calls/session
q
```
## cavega: another way to do this is to use the argument "lower.tail = FALSE" so you would do:
```{r}
ppois(8, lambda = 15, lower.tail = FALSE) #use lower.tail to tell the function not to include anything below 8 (so you don't have to do 1-p)
```

What is the probability that she will hear no calls in a session?
```{r}
dpois(x=0,lambda=15) #probability that she will hear 0 calls
```
## cavega: I did the same thing and got the same answer

What is the probability that she will hear exactly 3 calls in a session?
```{r}
dpois(x=3,lambda=15) #probability that she will hear 3 calls
```
## cavega: I did the same thing and got the same answer

Plot the relevant Poisson mass function over the values in range 0 ≤ x ≤ 30.
```{r}
x <- 0:30
l <- 15
barplot(dpois(x,l),names.arg=x,space=0,xlab="x",ylab="Pr(X = x)",main="Probability Mass Function")
```
## cavega: got the same result!

Simulate 104 results from this distribution (i.e., 2 years of Saturday monitoring sessions).
```{r}
set.seed(104)
pois <- rpois(104, lambda = 15)
pois
```
## cavega: This makes so much more sense. I attempted to use rnorm and did not get a normal distribution when I plotted the histogram.

Plot the simulated results using hist() and use xlim() to set the horizontal limits to be from 0 to 30. How does your histogram compare to the shape of the probability mass function you plotted above?
```{r}
hist(pois,xlim=c(0,30))
#The histogram above was more uniform/even/neat, which makes sense due to the variability of sampling. I was surprised that the peak of this histogram was 16 rather than 15. It was also interesting how there were no observations over 25, while the probability histogram above said there were some small chances of getting above 25. So even 104 observations wasn't enough to hit those small chance numbers. It looks like this histogram has more of a tail on the right (higher values) than the left (lower values). Overall, the bell shape is similar. 
```
## cavega: despite doing something different, my histogram is pretty similar to yours in that I had nothing below 10 and above 17. Clearly mine had more narrow margins than yours but I am not sure about the underlying reasons. 


#Five Challenges Reflection

1. Was a challenge to wrap my head around the fancy probability terminology (mass density probability, lambda, etc. etc.), got there eventually.
  ## cavega: Agreed. I was unsure what situations to use each function in.
  
2. Kept getting q & x mixed up for the poisson functions
  ## cavega: I avoided using them because I knew I would get messed up. Since it was relatively easy to remember q and x, I opted to not use them. 
  
3. For the first hw question, it would have been easier to find the probability of the event happening up to 8 calls instead of over 8 calls using cumulative probability. So I was trying to find a complicated way to get the probability of this, before I realized it was just 1 - the easy probability (if I'm even correct)
  ## cavega: we both got the same answer, but I included how I did it underneath your answer. I just looked in ?poisson and saw that lower.tail was an argument that can be used. 
  
4. I was also trying to write a complicated function for the last few hw questions before I googled & found there are random number generators for poisson distributions. 
  4a. I *think* I understand set.seed, but not 100% sure
    ## cavega: agreed. I really could not understand set.seed and then did not use it for the homework even though I should have
    
5. I do not fully understand what is happening with the automatic x axes that base R plots make before customizations, they suck</br>
  5a. why does ylim=0:1 work but xlim=0:30 doesn't work? 
  ## cavega: I had the same question! I would love to go over the differences between these plots in class. 
  
  ## cavega: Overall, your code was very readable and straightforward! Your annotations were great and insightful and when I compared answers it looks like we both ended up with similar results. Great job!
  